---
title: "Assignment 3, EDDA 2017"
author: "Martin de la Riva(11403799) and Kieran O'Driscoll(11426438), Group 23"
date: "24 April 2017"
output:
  pdf_document: default 
  html_document: default
highlight: tango
fontsize: 11pt
---

## Exercise 1
## Q1
```{r}
data = read.table("data/peruvians.txt", header=TRUE)
pairs(data[,-c(5,6,7)])
#labels = colnames(data)[-c(5,6,7)]
#counter = 1
#for(i in data[,-c(5,6,7)]) {
#  plot(i, migration, main=paste(labels[counter]," vs migration"), xlab=labels[counter], #ylab="Migration", pch=19)
#  abline(lm(migration~i), col="red") # regression line (y~x) 
#  counter = counter + 1 
#}
```
Based on the diagram above, age, weight and perhaps a case can be made for diastolic. These were chosen because their scatter plots show a cluster that shows the values are in porportion with migration whereas the othe graphs have a scatter plot that don't show any connection with migration.

##Q2
```{r}
 qqnorm(data[['length']])
```

Peruvians is not drawn from a normal distribution so therefore will use Spearmen correlation test to test for correlation.

```{r}
cor.test(age,migration,method="spearman")
```
The p-value recived is 0.002189 which falls under the 0.05 signicant level. Therefor the value is corrlated with migration. Since the value is quite low, it can be said there is a high positive correlation as the correlation increases as values get bigger.

```{r}
cor.test(length,migration,method="spearman")
```
With a P-value of 0.6087 it's clear that lenght and migration do not have a correlation.
```{r}
cor.test(wrist,migration,method="spearman")
```
Wrist gets a p-value of 0.1797 which means there is no correlation.
```{r}
cor.test(systolic,migration,method="spearman")
```
Systolic has no correlation with migration, p-value= 0.3054
```{r}
cor.test(diastolic,migration,method="spearman")
```
Diastolic has no correlation, the p-value was 0.6494
```{r}
cor.test(weight,migration,method="spearman")
```
Lastly, weight is correlated with migration having a p-value of 0.02861.
Two of the tree values predicted in question one were correlated, bith weight and age. 
##Exercise 1
##Q1

```{r, echo=FALSE}
data = read.table("data/run.txt", header=TRUE)
par(mfrow=c(1,2)) 
boxplot(before~drink,data=data, main="Energy Drink", 
  	xlab="Engery Drink", ylab="Perfomance Before")
boxplot(after~drink,data=data, main="Energy Drink", 
  	xlab="Engery Drink", ylab="Perfomance After")
#plot(i, migration, main=paste(labels[counter]," vs migration"), xlab=labels[counter], ylab="Migration", pch=19)
#abline(lm(migration~i), col="red") # regression line (y~x) 
```
From the boxplots it indicates that The engery actaully makes students worse after 60 mins and the lemonade makes students better.

```{r, echo=FALSE}
par(mfrow=c(2,2)) 
qqnorm(data[['before']]) 
qqnorm(data[['after']])
hist(data[['before']]) 
hist(data[['after']])

```
Data looks like it was drawn from a normal distribution.  With the exception of the histogram for "after" but there are only 12 entries per drink type which is a small amount to guage whether this is drawn from a normal distribution

##Q2
Using a two paired sample test
```{r, echo=FALSE}
filters = c('energy', 'lemo')
B=1000
value = vector("numeric", 2)
results = matrix(0,2,B)
par(mfrow=c(1,2))
counter = 1
for (i in filters) {
  filter = data['drink'] == i
  
  get_mean_diff=function(x,y) {mean(x-y)}#might need to change this
  result=numeric(B)
  for (j in 1:B)
  {
    new_data=t(apply(cbind(data['before'][filter],data['after'][filter]),1,sample))
    result[j]=get_mean_diff(new_data[,1],new_data[,2])
  }
  results[counter,] = result
  value[counter] = get_mean_diff(data['before'][filter],data['after'][filter])
  hist(results[counter,])
  counter = counter + 1
}

```

```{r}
  print("Energy")
  print(value[1])
  pl=sum(results[1,]<value[1])/B
  pr=sum(results[1,]>value[1])/B
  p=2*min(pl,pr)
  print(p)
  print("Lemo")
  print(value[2])
  pl=sum(results[2,]<value[2])/B
  pr=sum(results[2,]>value[2])/B
  p=2*min(pl,pr)
  print(p)
```
For both drinks there does not seem to be a signifcant diference between "before" and  "after". The p-values obtained are 0.128 and 0.4 for "energy" and "lemo" respectively.  

##Q3

Using permuation test permutation for independent samples.
 
```{r, echo=FALSE}
data['diff'] = data['after'] - data['before'] #absoulute
attach(data)
mystat=function(x) sum(residuals(x)^2)
B=1000
tstar=numeric(B)
for (i in 1:B)
{
  treatstar=sample(drink) ## permuting the labels
  tstar[i]=mystat(lm(diff~treatstar))
}
myt=mystat(lm(diff~drink))
hist(tstar)
myt
pl=sum(tstar<myt)/B
pr=sum(tstar>myt)/B

```
```{r, echo=FALSE}
print(pl)
```
The drink type does not affect the time difference from "before" to "after".

##Q4
There is only a small sample size, each drink gets allocated 12 people, it could be the case that one group were faster on average therefore would be faster before and after whereas some people might need more than 30 minutes to recover. Also, perhaps a bigger margin then 30 minutes for the gap between before and after. 

##Q5
It would have affected the time difference if people were still tired after the first run.

##Q6
?? Used paiered sample which doesn't need normal distribution.

```{r, echo=FALSE}

```

## Exercise 3
##Q1
```{r}
data = read.table("data/dogs.txt", header=TRUE)
par(mfrow=c(2,2))
boxplot(data,data=data)
qqnorm(data[['isofluorane']])
qqnorm(data[['halothane']])
qqnorm(data[['cyclopropane']])

#boxplot(after~drink,data=data, main="Energy Drink", 
#  	xlab="Engery Drink", ylab="Perfomance After")
```
Each drug type has 10 examples. All qqplots are close to a normal distribution with the exception of isofluorane which is displayed in top right. It can be presummed that by adding more data that the qqplots would more closely resembles a qqplot of a normal distubution. In this case it is reasonable to assume sample were taken from normal distribution. 


###Q2

```{r}
attach(data)
pvcaov=lm(isofluorane~halothane*cyclopropane,data=data)
anova(pvcaov)
```



### Q3

```{r}
kruskal.test(isofluorane,halothane)
kruskal.test(isofluorane,cyclopropane)
kruskal.test(cyclopropane,halothane)
```




