---
title: "Final Project, EDDA 2017"
author: "Martin de la Riva(11403799) and Kieran O'Driscoll(11426438), Group 23"
date: "01 June 2017"
output:
  pdf_document: default
  html_document: default
  word_document: default
highlight: tango
fontsize: 11pt
---

#Introduction
%Reserach questions
%Why is it important
%Whats is in the data

%Is a place a 

Air travel is increased by blah% in 2016 [reference] in the USA. However, what is relevant or significant when a crash occurs. In this paper this topic will be discussed. The first question that would be interesting is to check to see if the location of the crash matters. If a plane crashes more in a certain location this could  indicate a place where weather conditions are more hazardous, perhaps this was an era of conflict during that period of time or maybe its a popular air travel location. Also it would be of interest to see if the choice of Manfacturer has an impact on the number of fatalities.



  #provide refernece
  
#Data Distrubution

#Data findings

#Setting up experiment


```{r}
airline=read.csv("data/airline_data.csv", header=TRUE)
```

```{r}
temp = as.Date(airline$Date,'%m/%d/%Y')
airline['Date'] = format(temp,'%Y')

print(head(airline))
```

Given columns

```{r}
### Analysis of data
par(mfrow=c(1,3))
boxplot(airline$incidents_85_99, airline$incidents_00_14)
boxplot(airline$fatal_accidents_85_99, airline$fatal_accidents_00_14)
boxplot(airline$fatalities_85_99, airline$fatalities_00_14)

```