---
title: "Final Project, EDDA 2017"
author: "Martin de la Riva(11403799) and Kieran O'Driscoll(11426438), Group 23"
date: "01 June 2017"
output:
  pdf_document: default
  html_document: default
  word_document: default
highlight: tango
fontsize: 11pt
---

#Introduction
%Reserach questions
%Why is it important
%Whats is in the data

%Is a place a 

Air travel is increased by blah% in 2016 [reference] in the USA. However, what is relevant or significant when a crash occurs. In this paper this topic will be discussed. The first question that would be interesting is to check to see if the location of the crash matters. If a plane crashes more in a certain location this could  indicate a place where weather conditions are more hazardous, perhaps this was an era of conflict during that period of time or maybe its a popular air travel location. Also it would be of interest to see if the choice of Manfacturer has an impact on the number of fatalities.



  #provide refernece
  
#Data Distrubution

#Data findings

#Setting up experiment


```{r}
airline=read.csv("data/airline_data.csv", header=TRUE)
```

```{r}
temp = as.Date(airline$Date,'%m/%d/%Y')
airline['Date'] = format(temp,'%Y')
print(head(airline))
```

Given columns

```{r}
### Analysis of data
setup_data <- function(date, range, data) {
    airlines = unique(data$Operator)
    before = rep(0, length(airlines))
    after = rep(0, length(airlines))
    lower_bound = date - 4
    upper_bound = date + 5
    
    df <- data.frame(airlines, before, after)
    for(i in seq(from=1, to=length(airlines), by=1)){
      df$before[i] = sum(data$Fatalities[data$Date >= lower_bound & data$Date <= date & 
                                       data$Operator == df$airlines[i]])
      df$after[i] = sum(data$Fatalities[data$Date <= upper_bound & data$Date > date & 
                                       data$Operator == df$airlines[i]])
      #append(x, values, after=length(x))
    }

    #statements
    #return(something)
    return(df)
}

```