---
title: "Final Project, EDDA 2017"
author: "Martin de la Riva(11403799) and Kieran O'Driscoll(11426438), Group 23"
date: "01 June 2017"
output:
  pdf_document: default
  html_document: default
  word_document: default
highlight: tango
fontsize: 11pt
---

#Introduction
%Reserach questions
%Why is it important
%Whats is in the data

Air travel is increased by blah% in 2016 [reference] in the USA. Passenger Airlines that have a crash get an negative associtaion of not being safe. The first question the would be interesting it to check to see if an airline gets safer after a crash or incident? it is expecte dthat each crash has a impact on the Airline's stock market valuation as these valuations hugely rely on confidence. The question here is to see how much of an effect does a crash or an incident have on an airlines stock market evaluation.



  #provide refernece
  
#Data Distrubution

#Data findings

#Setting up experiment


```{r}
airline=read.csv("airline-safety.csv", header=TRUE)
```

```{r}
print(head(airline))
```

Given columns

```{r}
### Analysis of data
par(mfrow=c(1,3))
boxplot(airline$incidents_85_99, airline$incidents_00_14)
boxplot(airline$fatal_accidents_85_99, airline$fatal_accidents_00_14)
boxplot(airline$fatalities_85_99, airline$fatalities_00_14)

```