---
title: "Assignment 4 & 5, EDDA 2017"
author: "Martin de la Riva(11403799) and Kieran O'Driscoll(11426438), Group 23"
date: "08 May 2017"
output:
  word_document: default
  pdf_document: default
  html_document: default
highlight: tango
fontsize: 11pt
---

# Assignment 6

## Exercise 1
```{r}
fruitflies=read.table('fruitflies.txt', header=TRUE)
fruitflies$loglongevity <- log(fruitflies$longevity)
```

### 2.
```{r}
plot(loglongevity~thorax, pch=as.character(activity), col=as.numeric(activity), data=fruitflies)
abline(lm(loglongevity~thorax, data=fruitflies), col="blue")
```

In the plot above, longevity seems to increase with torax.

```{r}
# per activity:
boxplot(fruitflies$loglongevity~fruitflies$activity)
```

In these boxplots, it can be seen how high activity has a significantly low longevity than low and isolated activity. Those two last activities remain in similar longevity values.

```{r}
# thorax per activity (is there a significant difference between them?)
boxplot(fruitflies$thorax~fruitflies$activity)
```

Lastly, we plot activity with respect to thorax, to see if thorax lengths are equally distributed along the activity groups. This is important, as if one group was full of bigger thorax fruitflies, this could affect the result of the experiment. The activity groups are seen to actually have similar thorax values.

### 3.

```{r}
flieslm = lm(loglongevity~activity, data=fruitflies)
anova(flieslm)
```

Sexual activity seems to significantly influence longevity, with a p-value of 1.798e-07.

### 4.

```{r}
confint(flieslm)
```

In the Confidence Intervals for the different activities, this can be seen:
Confidence Intervals for high activity:   [3.4796296, 3.7246190]
Confidence Intervals for isolated - high: [0.3439909, 0.6904582]
Confidence Intervals for low - high:      [0.2244780, 0.5709453]
Therefore, high has the lowest longevity, and isolated the highest (isolated > low > high).

### 5.

```{r}
flieslm2 = lm(loglongevity~thorax+activity, data=fruitflies)
anova(flieslm2)
```
```{r}
drop1(flieslm2,test="F")
```

After testing it (both with anova and drop1, we get very similar results on both) we can conclude that both factors (thorax and activity) have a significant influence on the longevity.

### 6.

```{r}
summary(flieslm2)
```

As said before, the more sexual activity, the lowest longevity. Therefore, sexual activity influences negatively longevity (Isolated > Low > High).
final model: 1.21893 + 2.97899*thorax + 0.40998 (isolated)
                                      + 0.28570 (low)
                                      - 0.69568 (high)

For a fly with an average thorax:
```{r}
mean(fruitflies$thorax)
###    loglongevity_isolated= 1.21893 + 2.97899*0.8245333 + 0.40998 = 4.085186
1.21893 + 2.97899*mean(fruitflies$thorax) + 0.40998
###    loglongevity_low     = 1.21893 + 2.97899*0.8245333 + 0.28570 = 3.960906
1.21893 + 2.97899*mean(fruitflies$thorax) + 0.28570
###    loglongevity_high    = 1.21893 + 2.97899*0.8245333 - 0.69568 = 2.979526
1.21893 + 2.97899*mean(fruitflies$thorax) - 0.69568
```
```{r}
min(fruitflies$thorax)
###    loglongevity_isolated= 1.21893 + 2.97899*0.64 + 0.40998 = 3.535464
1.21893 + 2.97899*min(fruitflies$thorax) + 0.40998
###    loglongevity_low     = 1.21893 + 2.97899*0.64 + 0.28570 = 3.411184
1.21893 + 2.97899*min(fruitflies$thorax) + 0.28570
###    loglongevity_high    = 1.21893 + 2.97899*0.64 - 0.69568 = 2.429804
1.21893 + 2.97899*min(fruitflies$thorax) - 0.69568
```

### 7.

```{r}
plot(loglongevity~thorax, pch=unclass(activity), col=as.numeric(activity), data=fruitflies)
for (i in 1:3) abline(lm(loglongevity~thorax, data=fruitflies[as.numeric(fruitflies$activity)==i,]), col=i)
legend("bottomright", legend=levels(fruitflies$activity), col=c(1,2,3), lty=1, cex=0.5)
```

Lines have similar slope, which denotes that thorax influences similarly between activities.

### 8.

We prefer the analysis with thorax, as it is more complete. Also, it is possible to detect if there could be a difference in longevity between groups because of flies with significantly different thorax, instead of sexual activity. The analysis are fine, both factors do influence longevity.

### 9.

```{r}
qqnorm(residuals(flieslm2))
plot(fitted(flieslm2),residuals(flieslm2))
```

Normality is clearly met by the residuals, although heteroscedasticity is not that clear.

```{r}
flieslm3 = lm(longevity~thorax+activity, data=fruitflies)
qqnorm(residuals(flieslm3))
plot(fitted(flieslm3),residuals(flieslm3))
```

In this case heteroscedasticity seems more clear, which may mean that taking the logarithm of longevity was not a good decition.

## Exercise 3

### 1.
For this exercise  
```{r}
africa_data=read.table("africa.txt", header=TRUE)
par(mfrow=c(2,2))
hist(rpois(100, lambda = 1))
hist(rpois(100, lambda = 5))
hist(rpois(100, lambda = 10))
hist(rpois(100, lambda = 15))

```

```{r}
par(mfrow=c(2,2))
hist(rpois(1000, lambda = 1))
hist(rpois(1000, lambda = 5))
hist(rpois(1000, lambda = 10))
hist(rpois(1000, lambda = 15))
```

### 2.

```{r}
par(mfrow=c(2,2))
qqnorm(rpois(500, lambda = 1))
qqnorm(rpois(500, lambda = 15))
qqnorm(rpois(100, lambda = 15))
qqnorm(rpois(1000, lambda = 15))
```

What can be seen from question one is that as lambda and n increases, the poisson distribution further approximates a normal distribution. This is seen in the histgrams in question 1 and is further backed up by the QQ plots. As the poisson distribution gets closer to approximating the normal distribution, the Poisson distributions would be in the same location-scale family as they would have properties similar to a normal distribution.


#transform distrubution, see if it looks same afterwards
```{r}
africa_data=read.table("africa.txt", header=TRUE)
par(mfrow=c(3,2))
hist(rpois(1000, lambda = 2))
hist(rpois(1000, lambda = 25))
hist(rpois(1000, lambda = 50))
hist(rpois(1000, lambda = 75))
hist(rpois(1000, lambda = 100))
hist(rpois(1000, lambda = 240))
```


### 3

```{r}
#Analysisof variance
galaglm_full=glm(miltcoup~oligarchy+pollib+parties+pctvote+popn+size+numelec +numregim,family=poisson,data=africa_data)
error1 = summary(galaglm_full)$r.squared
summary(galaglm_full)
```

### 4.
Removing numelec has it had the gighest p-value.
```{r}

#Analysisof variance
galaglm=glm(miltcoup~oligarchy+pollib+parties+pctvote+popn+size +numregim,family=poisson,data=africa_data)
summary(galaglm)
```
Removing numregim.

```{r}
galaglm=glm(miltcoup~oligarchy+pollib+parties+pctvote+popn+size,family=poisson,data=africa_data)
summary(galaglm)
```
Removing size.

```{r}
galaglm=glm(miltcoup~oligarchy+pollib+parties+pctvote+popn,family=poisson,data=africa_data)
summary(galaglm)
```
Removing popn.

```{r}
galaglm=glm(miltcoup~oligarchy+pollib+parties+pctvote,family=poisson,data=africa_data)
summary(galaglm)
```
And finally removing pctvote.

```{r}
galaglm=glm(miltcoup~oligarchy+pollib+parties,family=poisson,data=africa_data)
summary(galaglm)
```


###5.

```{r}
par(mfrow=c(2,2))
attach(africa_data)
plot(fitted(galaglm),residuals(galaglm))
plot(residuals(galaglm),parties)
plot(residuals(galaglm),pollib)
plot(residuals(galaglm),oligarchy)
```

```{r}
par(mfrow=c(1,3))
plot(residuals(galaglm_full),parties)
plot(log(fitted(galaglm)),residuals(galaglm))
plot(log(fitted(galaglm)),residuals(galaglm,type="response"))
#plot(1:36,cooks.distance(galaglm))
print(residuals(galaglm))
```


The response residuals clearly increase with the (logarithm) of the fitted values,
as expected under a Poisson model.

Next, the model from Q3 will be investigated to see if it follows the same pattern.

```{r}
par(mfrow=c(1,3))
plot(residuals(galaglm_full),parties)
plot(log(fitted(galaglm_full)),residuals(galaglm_full))
plot(log(fitted(galaglm_full)),residuals(galaglm_full,type="response"))
print(residuals(galaglm_full))
```

While the response residuals clearly increase with the (logarithm) of the fitted values, the step down approach looks to contain less clear patterns in it's data. In the full model with all explanatory variables, there looks to be much clearer linear structures within the data compared to the step down model. This could be due to the removed variables. During the process of the step down approach in Q4 what can be seen is that some of the variables become less significant as variables are removed. For example, in the last step the variable pctvote is removed. This reduces the p value of the variable parties. By removing pctvote, the relationship between pctvote and parties is removed which will affect the model.

```{r}
pairs(pctvote~parties)
```

However, from the pairs graph there does not seem to be a correlation.